---
title: "Installing the Tidyverse on Linux TSD"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{tidyverse-tsd-install}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The tidyverse meta-package has become a well-used suite of packages for data handling and visualisation.
However, the tidyverse has some system and package dependencies that makes it tricky to install on Linux TSD. 
Mainly because it depends on the [stringi](https://cran.r-project.org/web/packages/stringi/index.html) package, which needs internet access to install text information that cannot be stored within the main package itself. 

## Pre-installed bundles
There are three modules on TSD where it have bundled popular packages for you within the R-package it self.
If you are happy using the modules, you can use any of the "foss" modules, like `R/4.0.0-foss-2020a`. 

```{bash}
# example 
module load R/4.0.0-foss-2020a

# Then start R as you would normally
```


## Prepare the tidyverse packages for install
Firstly, prepare all the necessary packages and dependencies for install on TSD using the `tsd_package_prepate()` function.

```{r setup}
library(tsdrtools)
tsd_package_prepare("tidyverse")
```

All the tidyverse packages and their dependencies will now download into a file called `tidyverse.zip`. 
Import this file into your TSD project.

## Prepare your TSD environment for tidyverse

### Linux OS libs
Several of the tidyverse packages require your linux system to have certain OS libraries installed for them to work. 
You will need to contact USIT with information on which such system libraries you require, and which project to have them installed for (we recommend installing for the entire project, rather than individual users).

Needed OS libs:

- libssh2-devel  
- libssl-dev  
- libxml2-devel 
- libcurl-devel
- harfbuzz-devel
- libpng-devel (optional, but recommended)  

## Install
Once all the libs are installed by IT, and you load gcc before starting R, you are ready to get tidyverse installed.

```{r}
tsd_package_prepare("path/to/tidyverse.zip")
```

This should now hopefully install without errors. 
If you still cannot install tidyverse with these steps, report an [issue](https://github.com/LCBC-UiO/tsdrtools/issues).
If you know of additional needed steps for successful install, either do a direct PR for adding that information to this document, or , report an [issue](https://github.com/LCBC-UiO/tsdrtools/issues).
We'd love feedback to improve the instructions!


